---
title: ""
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    css: !expr here::here("www", "web_report.css")
    editor_options:
      chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />



```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
options(scipen = 999)


```



# `gt` table theme examples

This contains brief examples of tables in the Urban format.
<br>

```{r}
library(tidyverse)
library(gt)

df <- tribble(
  ~`City or MSA`, ~`Home price`, ~`Fair market rent ($ millions)`, ~`Rent`, ~`First month principal`, ~`Rent vouchers`, ~`Homeownership vouchers`,
  "Houston, TX", 589, 589, 589, 589, 589, 589,
  "Philadelphia, PA", 296, 296, 296, 296, 296, 296,
  "Columbus, OH", 883, 883, 883, 883, NA_integer_, 883,
  "El Paso, TX", 296, 296, 296, 296, 296, 296,
  "Louisville, KY", 883, 883, 883, 883, 883, 883
)




create_table <- function(data_df, 
                         rowname_col = NA,
                         fig_num = "",
                         title_text = ""){
  # random_id = random_id(n=10)

  random_id = "urban_table"
  
  basic_table = data_df %>% 
    gt(id = random_id, rowname_col = rowname_col) %>% 
    tab_options(table.width = px(760),
                table.align = "left", 
                heading.align = "left",
                # TODO: Discuss with Comms whether border should extend across 
                # whole row at bottom or just across data cells
                table.border.top.style = "hidden",
                table.border.bottom.style = "transparent",
                heading.border.bottom.style = "hidden",
                # Need to set this to transparent so that cells_borders of the cells can display properly and 
                table_body.border.bottom.style = "transparent",
                table_body.border.top.style = "transparent",
                # column_labels.border.bottom.style = "transparent",
                column_labels.border.bottom.width = px(1),
                column_labels.border.bottom.color = "black",
                # row_group.border.top.style = "hidden",
                # Set font sizes
                heading.title.font.size = px(14),
                heading.subtitle.font.size = px(18),
                column_labels.font.size = px(14),
                table.font.size = px(14),
                source_notes.font.size = px(12),
                footnotes.font.size = px(12),
                # Set row group label and border options
                row_group.font.size = px(14),
                row_group.border.top.style = "transparent",
                row_group.border.bottom.style = "hidden",
                stub.border.style = "dashed",

                ) %>% 
    tab_header(
      title = fig_num,# "eyebrow",
      subtitle = title_text) %>%  #"Top 10 Banks (by Dollar Volume) for Community Development Lending") %>% 
    # Bold title, subtitle, and columns
    tab_style(
      style = cell_text(color = "black", weight = "bold", align = "left"),
      locations = cells_title("subtitle")
    ) %>% 
    tab_style(
      style = cell_text(color = "#696969", weight = "normal", align = "left", transform = "uppercase"),
      locations = cells_title("title")
    ) %>% 
    tab_style(
      style = cell_text(color = "black", weight = "bold", size = px(14)),
      locations = cells_column_labels(gt::everything())
    ) %>% 
    # Italicize row group and column spanner text
    tab_style(
      style = cell_text(color = "black", style = "italic", size  = px(14)),
      locations = gt::cells_row_groups()
    ) %>% 
    tab_style(
      style = cell_text(color = "black", style = "italic", size  = px(14)),
      locations = gt::cells_column_spanners()
    ) %>% 
    opt_table_font(
        font = list(
          google_font("Lato"),
          default_fonts()
        )
      ) %>% 
    # Adjust cell borders for all cells, small grey bottom border, no top border
    tab_style(
      style = list(
        cell_borders(
          sides = c("bottom"),
          color = "#d2d2d2",
          weight = px(1)
        )
      ),
      locations = list(
        cells_body(
          columns =  gt::everything()
          # rows = gt::everything()
        )
      )
    )  %>%
    tab_style(
      style = list(
        cell_borders(
          sides = c("top"),
          color = "#d2d2d2",
          weight = px(0)
        )
      ),
      locations = list(
        cells_body(
          columns =  gt::everything()
          # rows = gt::everything()
        )
      )
    )  %>%
    # Set missing value defaults
    fmt_missing(columns = gt::everything(), missing_text = "—") %>%
    # Set css for all the things we can't finetune exactly in gt, mostly t/r/b/l padding
    opt_css(
      css = str_glue("
      #{random_id} .gt_row {{
        padding: 15px 15px 15px 15px;
      }}
      #{random_id} .gt_sourcenote {{
        padding: 16px 0px 0px 0px;
      }}
      #{random_id} .gt_footnote {{
        padding: 16px 0px 0px 0px;
      }}
      #{random_id} .gt_subtitle {{
        padding: 0px 0px 22px 0px;
      }}
      #{random_id} .gt_col_heading {{
        padding: 10px 15px 10px 15px;
      }}
      #{random_id} .gt_col_headings {{
        padding: 0px 0px 0px 0px;
        border-top-width: 0px;
      }}
      #{random_id} .gt_group_heading {{
        padding: 15px 0px 0px 0px;
      }}
      #{random_id} .gt_stub {{
        border-bottom-width: 1px;
        border-bottom-style: solid;
        border-bottom-color: #d2d2d2;
        border-top-color: black;
        text-align: left;
      }}
      #{random_id} .gt_grand_summary_row {{
        border-bottom-width: 1px;
        border-top-width: 1px;
        border-bottom-style: solid;
        border-bottom-color: #d2d2d2;
        border-top-color: #d2d2d2;
      }}
      #{random_id} .gt_summary_row {{
        border-bottom-width: 1px;
        border-top-width: 1px;
        border-bottom-style: solid;
        border-bottom-color: #d2d2d2;
      }}
      #{random_id} .gt_column_spanner {{
        padding-top: 10px;
        padding-bottom: 10px;
      }}


      ") %>% as.character()
    )
  
  return(basic_table)
}

  
  
```


## Basic table (no customization)

This is the table as it would look if someone just applied the Urban theme to it (ie one line of code).

```{r reprex_for_github_issue}

basic_table = create_table(df,
                           fig_num = "eyebrow",
                           title_text = "Top 10 Banks (by Dollar Volume) for Community Development Lending" )
basic_table
```

<br>
<br>


## Basic table with customization (widths, column capitalization)

This is the table if someone added a few more customizations for column widths, and column name spacing/selective capitalization (ie uncapitalizing `$ millions` in the column title and put it on a new line). This will require a bit more education and manual work on the programmers side (Click the code button to see how much work) but is a better looking table.

```{r}
 


add_housing_table_specific_customizations = function(gt_object){
  gt_object %>% 
  cols_width(
    "City or MSA" ~ px(132),
    "Home price" ~ px(108),
    "Fair market rent ($ millions)" ~ px(108),
    "Rent" ~ px(90),
    "First month principal" ~ px(108),
    "Rent vouchers" ~ px(90),
    "Homeownership vouchers" ~ px(124),
  ) %>% 
  cols_label(
    `Fair market rent ($ millions)` = html('<b>Fair market rent</b><br><span style="font-weight:normal">($ millions)</span>'),
    `Homeownership vouchers` = html("Homeowner-<br>ship vouchers")
     # `Fair market rent ($ millions)` = md("**Fair market rent**($ millions)"),
  )
}

basic_table %>%
  add_housing_table_specific_customizations()

```


<br>
<br>

## Sourcenotes
```{r}
basic_table %>% 
  add_housing_table_specific_customizations() %>% 
  tab_source_note(source_note = "Source: Narayanan, and Williams, 2019") 

```

<br>
<br>

## Footnotes
```{r}
basic_table %>% 
  add_housing_table_specific_customizations() %>% 
  tab_footnote(
    footnote = "Indexed to 2006 Dollars", 
    locations = cells_column_labels(
      columns = vars("Home price") # note
      ))
```



<br>
<br>

## Row group labels

```{r}
df_grouped <- df %>% 
  mutate(groups = c("Big Cities", "Big Cities", "Smaller Cities", "Smaller Cities", "Smaller Cities")) %>% 
  group_by(groups)


basic_table_row_group_labels = create_table(df_grouped,
                          fig_num = "eyebrow",
                          title_text = "Top 10 Banks (by Dollar Volume) for Community Development Lending"
                          ) %>% 
  add_housing_table_specific_customizations()

basic_table_row_group_labels

```

<br>
<br>

## Spanner column labels (1) 

```{r}
basic_table %>% 
  add_housing_table_specific_customizations() %>% 
  tab_spanner(
    label = "Names",
    columns = 1:2,
  ) %>% 
  tab_style(
    style = cell_text(color = "black", style = "italic", size  = px(14)),
    locations = gt::cells_column_spanners()
  )

```


## Spanner column labels (2) 

```{r}
basic_table %>% 
  add_housing_table_specific_customizations() %>% 
  tab_spanner(
    label = "Names",
    columns = 1:2,
  ) %>% 
  tab_spanner(
    label = "Dollars",
    columns = 3:4
  ) %>% 
  tab_style(
    style = cell_text(color = "black", style = "italic", size  = px(14)),
    locations = gt::cells_column_spanners()
  )

```


<br>
<br>


## Grand summary rows 
```{r}
create_table(df, 
             rowname_col = "City or MSA",
             fig_num = "eyebrow",
             title_text = "Top 10 Banks (by Dollar\ Volume) for Community Development Lending" ) %>% 
  add_housing_table_specific_customizations() %>% 
  grand_summary_rows(
    columns = 2:4,
    fns = list(
      # add summary stats
      Total = ~sum(.)
    )
  ) 


```




<br>
<br>

## Everything in a messy table
```{r}
create_table(df_grouped, 
             rowname_col = "City or MSA",
             fig_num = "eyebrow",
             title_text = "Top 10 Banks (by Dollar Volume) for Community Development Lending" ) %>% 
  add_housing_table_specific_customizations() %>% 
  summary_rows(
    groups = TRUE,
    columns = 2:4,
    fns = list(
      # add summary stats
      Total = ~sum(.)
    )
  ) %>% 
  tab_source_note(source_note = "Source: Narayanan, and Williams, 2019") %>% 
  tab_spanner(
    label = "Names",
    columns = 1:2,
  ) %>% 
  tab_spanner(
    label = "Dollars",
    columns = 3:4
  ) %>% 
  tab_footnote(
    footnote = "Indexed to 2006 Dollars", 
    locations = cells_column_labels(
      columns = vars("Home price") # note
    )) %>% 
  tab_style(
    style = cell_text(color = "black", style = "italic", size  = px(14)),
    locations = gt::cells_column_spanners()
  )


```



<br>
<br>
<br>
<br>
<br>
<br>

